---
title: "Practical Machine Learning Course Project"
author: "Jason Weng"
highlighter: highlight.js
output:
  html_document: default
  pdf_document: default
mode: selfcontained
hitheme: tomorrow
subtitle: Exercise Manner Prediction
framework: io2012
widgets: mathjax
---

````{r setup, cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE, results='hide', error=FALSE}
# make this an external chunk that can be included in any file
library(knitr)
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')

#options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
#knit_hooks$set(plot = knitr:::hook_plot_html)
runif(1)
```

## Executive Summary 
* Download the data.
* Preprocess the data and choose predictiont varaibles. 
* Split the training data to 60% to 40% to valide the model using random forest. 
* Get a modle with good performance and appply to final testing set.

## Download traing data
```{r fig.height=4,fig.width=4,echo=TRUE}
require(data.table)
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training.dt <- fread(url)
str(training.dt)

url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing.dt <- fread(url)
```

## Process Training Data
### Find predictors: Choose variables contains(Belt, arm, dumbbell, and forearm ) and do not have any missing value as predictors
```{r fig.height=4,fig.width=4,echo=TRUE}
is.missing <- sapply(testing.dt, function (x) any(is.na(x) | x == ""))
is.predictor <- !is.missing & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(testing.dt))
pred.candidates <- names(testing.dt)[is.predictor]
pred.candidates
```

### Build the training data with predictors and target.
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
train.vars <- c("classe", pred.candidates)
train.dt <- training.dt[, train.vars, with=FALSE]
dim(train.dt)
```

### Make "classe" variable into a factor 
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
train.dt <- train.dt[, classe := factor(train.dt[, classe])]
train.dt[, .N, classe]
```


### Split the dataset into a 60% training and 40% probing dataset
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
require(caret)
seed <- as.numeric(as.Date("2017-06-14"))
set.seed(seed)
in.train <- createDataPartition(train.dt$classe, p=0.6)
trainset <- train.dt[in.train[[1]]]
testset <- train.dt[-in.train[[1]]]
```

### Preprocess the data.
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
pred.data <- trainset[, pred.candidates, with=FALSE]
pre.proc <- preProcess(pred.data)
print (pre.proc)
train.proc <- predict(pre.proc, pred.data)
train.input <- data.table(data.frame(classe = trainset[, classe], train.proc))

test.data <- testset[, pred.candidates, with=FALSE]
test.proc <- predict(pre.proc, test.data)
test.input <- data.table(data.frame(classe = testset[, classe], test.proc))

nzv <- nearZeroVar(train.input, saveMetrics=TRUE)
if (any(nzv$nzv)) nzv else message("No variables with near zero variance")
```


## Explore the data for prediction
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
histGroup <- function (data, regex) {
  col <- grep(regex, names(data))
  col <- c(col, which(names(data) == "classe"))
  require(reshape2)
  n <- nrow(data)
  DMelted <- melt(data[, col, with=FALSE][, rownum := seq(1, n)], id.vars=c("rownum", "classe"))
  require(ggplot2)
  ggplot(DMelted, aes(x=classe, y=value)) +
    geom_violin(aes(color=classe, fill=classe), alpha=1/2) +
    facet_wrap(~ variable, scale="free_y") +
    scale_color_brewer(palette="Spectral") +
    scale_fill_brewer(palette="Spectral") +
    labs(x="", y="") +
    theme(legend.position="none")
}
histGroup(train.input, "belt")
histGroup(train.input, "[^(fore)]arm")
histGroup(train.input, "dumbbell")
histGroup(train.input, "forearm")
```


## Traing the model
### Set up clusters
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
require(parallel)
require(doParallel)
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
```
### Set the control parameters.
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
ctrl <- trainControl(classProbs=TRUE,
                     savePredictions=TRUE,
                     allowParallel=TRUE)
```
### Fit Model
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
method <- "rf"
system.time(trainingModel <- train(classe ~ ., data=train.input, method=method))
stopCluster(cl)
```

### Evaluate the model on the training set
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
trainingModel
hat <- predict(trainingModel, train.input)
confusionMatrix(hat, train.input[, classe])
```

### Evaluate the model on the testing set
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
hat <- predict(trainingModel, test.input)
confusionMatrix(hat, test.input[, classe])
```

### Save the model
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
varImp(trainingModel)
trainingModel$finalModel
save(trainingModel, file="trainingModel.RData")
```
#### Error rate is less than 1% 

## Preidict on the test set 
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
load(file="trainingModel.RData", verbose=TRUE)
#url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#testing.dt <- fread(url)


final.input <- predict(pre.proc, testing.dt[, pred.candidates, with=FALSE])
hat <- predict(trainingModel, final.input)
testing.dt <- cbind(hat , testing.dt)
subset(testing.dt, select=names(testing.dt)[grep("belt|[^(fore)]arm|dumbbell|forearm", names(testing.dt), invert=TRUE)])
```
## Save to target format
```{r fig.height=4,fig.width=6,fig.align='center',echo=TRUE}
write.files = function(x){
  n = length(x)
  path <- "predictionAssignment_files/answers"
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=file.path(path, filename),quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
write.files(hat)
```


